Proof-Reconstruction in Type Theory for Propositional Logic

### 2017-10-22 paper/sections/proof-reconstruction/strip.tex
[ paper ] minor fixes.
[ paper ] fixing lemmas.
last changes.

We now define the $\fstrip$ function in Eq.~\ref{eq:strip-unfixed} that
yields the conjunction of subgoals that implies the goal of the problem in
the \Metis \TSTP derivations.

\begin{equation}
\label{eq:strip-unfixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip &(φ₁ ∧ φ₂)     &= \fuh~(\fstrip~φ₁) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂)\\
\fstrip &(φ₁ ∨ φ₂)     &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂)\\
\fstrip &(φ₁ ⇒ φ₂)     &= \fuh~(φ₁ ⇒ \fstrip~φ₂)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) &= \fuh~(\fstrip~¬ φ₁) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) &= \fuh~(\fstrip~φ₁) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂)\\
\fstrip &(¬ ¬ φ₁)      &= \fuh~(\fstrip~φ₁)\\
\fstrip &(¬ ⊥)         &= ⊤\\
\fstrip &(¬ ⊤)         &= ⊥\\
\fstrip &φ             &= φ
\end{array}
\end{aligned}
\end{equation}

@@

\begin{equation}
\label{eq:strip-fixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip &(φ₁ ∧ φ₂)     ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     ~(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      ~(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         ~(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         ~(\suc~n) &= ⊥\\
\fstrip &φ             ~\emph{n} &= φ
\end{array}
\end{aligned}
\end{equation}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $n : \Nat$ the complexity measure of the strip function in
Eq.~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fstrip~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

@@

Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip~(¬¬φ₁)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
    \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
\RightLabel{by def.}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n)$}
\RightLabel{lemma~\ref{lem:inv-uh-lem}}
\UnaryInfC{$Γ ⊢ \fstrip~¬ φ₁~n$}
\RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬φ₁$}
\end{bprooftree}
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip~¬ φ₂~n$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~¬ φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}
\end{sketchproof} % the proof continue...

@@

\begin{proof}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$ Γ ⊢ s_1 $}
  \AxiomC{$ Γ ⊢ s_2 $}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ s_1\wedge s_2$}
  \AxiomC{$\fstrip~φ~n ≡ s_1\wedge s_2 $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \fstrip~φ~n$}
  \RightLabel{lemma~\ref{lem:lem-inv-strip}}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}

>>>>

We now define the $\fstrip_0$ function in Eq.~\ref{eq:strip-unfixed} that
yields the conjunction of subgoals that implies the goal of the problem in
the \Metis \TSTP derivations.

\begin{equation}
\label{eq:strip-unfixed}
\begin{aligned}
&\hspace{.495mm}\fstrip_0 : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip_0 &(φ₁ ∧ φ₂)     &= \fuh~(\fstrip_0~φ₁) ∧ \fuh~(φ₁ ⇒ \fstrip_0~φ₂)\\
\fstrip_0 &(φ₁ ∨ φ₂)     &= \fuh~(¬ φ₁ ⇒ \fstrip_0~φ₂)\\
\fstrip_0 &(φ₁ ⇒ φ₂)     &= \fuh~(φ₁ ⇒ \fstrip_0~φ₂)\\
\fstrip_0 &(¬ (φ₁ ∧ φ₂)) &= \fuh~(φ₁ ⇒ \fstrip_0~¬ φ₂)\\
\fstrip_0 &(¬ (φ₁ ∨ φ₂)) &= \fuh~(\fstrip_0~¬ φ₁) ∧ \fuh~(¬ φ₁ ⇒ \fstrip_0~¬ φ₂)\\
\fstrip_0 &(¬ (φ₁ ⇒ φ₂)) &= \fuh~(\fstrip_0~φ₁) ∧ \fuh~(φ₁ ⇒ \fstrip_0~¬ φ₂)\\
\fstrip_0 &(¬ ¬ φ₁)      &= \fuh~(\fstrip_0~φ₁)\\
\fstrip_0 &(¬ ⊥)         &= ⊤\\
\fstrip_0 &(¬ ⊤)         &= ⊥\\
\fstrip_0 &φ             &= φ
\end{array}
\end{aligned}
\end{equation}

@@

%  1137  git add .
%  1138  git commit -am "[ paper ] strip proof."
%  1139  git pull origin paper
%  1140  git push origin paper
\begin{equation}
\label{eq:strip-fixed}
\begin{aligned}
&\hspace{.495mm}\fstrip : \Prop → \Nat → \Prop\\
&\begin{array}{lll}
\fstrip &(φ₁ ∧ φ₂)     ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ∨ φ₂)     ~(\suc~n) &= \fuh~(¬ φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(φ₁ ⇒ φ₂)     ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~φ₂~n)\\
\fstrip &(¬ (φ₁ ∧ φ₂)) ~(\suc~n) &= \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ∨ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ (φ₁ ⇒ φ₂)) ~(\suc~n) &= \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~¬ φ₂~n)\\
\fstrip &(¬ ¬ φ₁)      ~(\suc~n) &= \fuh~(\fstrip~φ₁~n)\\
\fstrip &(¬ ⊥)         ~(\suc~n) &= ⊤\\
\fstrip &(¬ ⊤)         ~(\suc~n) &= ⊥\\
\fstrip &φ             ~\text{n} &= φ
\end{array}
\end{aligned}
\end{equation}

\begin{mainlemma}
\label{lem:lem-inv-strip}
Let $\text{n} : \Nat$ the complexity measure of the strip function in
Eq.~(\ref{eq:strip-unfixed}).
If $Γ ⊢ \fstrip~φ~n$ then $Γ ⊢ φ$.
\end{mainlemma}

@@

Case $φ ≡ ¬ ¬ φ₁$.
\begin{equation*}
  \begin{bprooftree}
 \AxiomC{$Γ ⊢ \fstrip~(¬¬φ₁)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
  \begin{bprooftree}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{∧-proj₁}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ \fstrip~φ₁~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₁$}
    \end{bprooftree}
\end{equation*}

\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{Γ ⊢ φ₁}
  \AxiomC{$Γ ⊢ \fstrip~(φ₁ ∧ φ₂)~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~φ₁~n) ∧ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{$∧$-proj$₂$}
  \UnaryInfC{$Γ ⊢ \fuh~(φ₁ ⇒ \fstrip~φ₂~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ φ₁ ⇒ \fstrip~φ₂~n$}
  \RightLabel{$⇒$-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ φ₂$}
  \end{bprooftree}
\end{equation*}

@@

\begin{equation*}
(\mathcal{D})\hspace{3mm}
\begin{bprooftree}
\AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
\RightLabel{by def.}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
\RightLabel{∧-proj₁}
\UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n)$}
\RightLabel{Lemma~\ref{lem:inv-uh-lem}}
\UnaryInfC{$Γ ⊢ \fstrip~¬ φ₁~n$}
\RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬φ₁$}
\end{bprooftree}
\end{equation*}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$\mathcal{D}$}
  \UnaryInfC{$Γ ⊢ ¬ φ₁$}
  \AxiomC{$Γ ⊢ \fstrip~(¬ (φ₁ ∨ φ₂))~(\suc~n)$}
  \RightLabel{by def.}
  \UnaryInfC{$Γ ⊢ \fuh~(\fstrip~¬ φ₁~n) ∧ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{∧-proj₂}
  \UnaryInfC{$Γ ⊢ \fuh~(¬ φ₁ ⇒ \fstrip~¬ φ₂~n)$}
  \RightLabel{Lemma~\ref{lem:inv-uh-lem}}
  \UnaryInfC{$Γ ⊢ ¬ φ₁ ⇒ \fstrip~¬ φ₂~n$}
  \RightLabel{⇒-elim}
  \BinaryInfC{$Γ ⊢ \fstrip~¬ φ₂~n$}
  \RightLabel{by hip.}
  \UnaryInfC{$Γ ⊢ ¬ φ₂$}
  \end{bprooftree}
\end{equation*}
\end{sketchproof} % the proof continue...

@@

\begin{proof}
\begin{equation*}
  \begin{bprooftree}
  \AxiomC{$ Γ ⊢ s_1 $}
  \AxiomC{$ Γ ⊢ s_2 $}
  \RightLabel{∧-intro}
  \BinaryInfC{$Γ ⊢ s_1\wedge s_2$}
  \AxiomC{$\fstrip~φ~n ≡ s_1\wedge s_2 $}
  \RightLabel{subst}
  \BinaryInfC{$Γ ⊢ \fstrip~φ~n$}
  \RightLabel{Lemma~\ref{lem:lem-inv-strip}}
  \UnaryInfC{$Γ ⊢ φ$}
\end{bprooftree}
\end{equation*}
\end{proof}
